---
title: "chetah_tirosh"
author: "Shaila Fye"
date: "7/28/2020"
output: html_document
---
TIROSH data --- /Users/shailafye/Documents/summer2020/tirosh/melanoma_tirosh_alldata

https://www.bioconductor.org/packages/release/bioc/vignettes/CHETAH/inst/doc/CHETAH_introduction.html#sce
https://github.com/jdekanter/CHETAH (ReadMe)
https://nbisweden.github.io/excelerate-scRNAseq/session-celltypeid/celltypeid.html

Make a single cell experiment object with reference list and data


```{r}
## Install BiocManager is neccesary
if (!require("BiocManager")) {
    install.packages("BiocManager")
}
BiocManager::install('CHETAH')

# Load the package
library(CHETAH)
```



A SingleCellExperiment holds three things:
counts: assays -- as a list of Matrices
meta-data: colData -- as DataFrames
reduced dimensions (e.g. t-SNE, PCA): ReducedDims -- as a SimpleList of 2-column data.frames or matrices

CHETAH needs
a reference SingleCellExperiment with: an assay, a colData column with the corresponding cell types (default “celltypes”)
an input SingleCellExperiment with:an assay, a reducedDim (e.g. t-SNE)

```{r}
myData <- read.delim(file="/Users/shailafye/Documents/summer2020/tirosh/melanoma_tirosh_alldata", comment = "#", stringsAsFactors=F)
#tsne_hn <- readRDS(file = "/Users/shailafye/Documents/summer2020/puram/tnse_attempt1.rds")
library(SingleCellExperiment)

sce <- SingleCellExperiment(list(counts=as.matrix(myData)))
head(sce)
cell.labels <- colnames(myData)

sc <- SingleCellExperiment(
    list(counts=as.matrix(myData)),
    colData=DataFrame(label=cell.labels),
    rowData = DataFrame(length=rownames(myData)),
    metadata=list(study="GSE103322"))



```


Reference dataset
/Users/shailafye/Documents/summer2020/puram/chetah/CHETAH_TME_reference.Rdata
https://figshare.com/s/aaf026376912366f81b6

```{r}
load("/Users/shailafye/Documents/summer2020/puram/chetah/CHETAH_TME_reference.Rdata")
ls() 
#reference --> variable stored
head(reference)
reff<-read.delim(file = "/Users/shailafye/Documents/summer2020/puram/chetah/ribosomal.txt")




```



run CHETAH
```{r}
input <- CHETAHclassifier(input = sc, ref_cells = reference) #,thresh = (default = 0.1) can change 
#takes 1 min to run
head(input)

class(input)
```


input$celltype_CHETAH
a named character vector that can directly be used in any other workflow/method.
“hidden” int_colData and int_metadata, not meant for direct interaction, but
which can all be viewed and interacted with using: PlotCHETAH and CHETAHshiny

```{r}
input$celltype_CHETAH
table(input$celltype_CHETAH)
input$celltype_CHETAH
class(input$celltype_CHETAH)
x <- input$celltype_CHETAH
as.list(x)
write.table(input$celltype_CHETAH,file="/Users/shailafye/Documents/summer2020/tirosh/tirosh_celltype_CHETAH_1_barcodes", sep = "\t")



#manipulate output file to look like the others (without cell barcode)
#numbered

cell_chetah <- as.data.frame(input$celltype_CHETAH)
row.names(cell_chetah) <- NULL
colnames(cell_chetah) <- NULL
table(input$celltype_CHETAH)
cell_chetah
colnames(cell_chetah) <- "x"
write.table(cell_chetah,file="/Users/shailafye/Documents/summer2020/tirosh/tirosh_celltype_CHETAH_1", sep = "\t")

```



CHETAH constructs a classification tree by hierarchically clustering the reference data. The classification is guided by this tree. In each node of the tree, input cells are either assigned to the right, or the left branch. A confidence score is calculated for each of these assignments. When the confidence score for an assignment is lower than the threshold (default = 0.1), the classification for that cell stops in that node.

This results in two types of classifications:

final types: Cells that are classified to one of the leaf nodes of the tree (i.e. a cell type of the reference).
intermediate types: Cells that had a confidence score lower than the threshold in a certain node are assigned to that intermediate node of the tree. This happens when a cell has approximately the same similarity to the cell types in right and the left branch of that node.
CHETAH generates generic names for the intermediate types: “Unassigned” for cells that are classified to the very first node, and “Node1”, “Node2”, etc for the additional nodes

```{r}
#To plot the final types:
PlotCHETAH(input = input_mel)

#Conversely, to color the intermediate types:
PlotCHETAH(input = input_mel, interm = TRUE)



```

If you would like to use the classification, and thus the colors, in another package (e.g. Seurat2), you can extract the colors using:

colors <- PlotCHETAH(input = input_mel, return_col = TRUE)



#looking at reference dataset information
```{r}
load(file ="/Users/shailafye/Documents/summer2020/puram/chetah/headneck_ref.rda")
ls() #"headneck_ref" 
headneck_ref
str(headneck_ref)
head(headneck_ref)
headneck_ref$celltypes
```


